---
title: "Simulated benchmarking of hashing based algorithms"
author: "Michael Lynch"
date: "`r Sys.Date()`"
bibliography: references.bib
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulated benchmarking of hashing based algorithms}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options: 
  markdown: 
    wrap: 80
---

```{r eval=TRUE, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    message = FALSE,
    fig.width = 7.5
)
```

```{r message=FALSE, warning=FALSE}
library(ComplexHeatmap)
library(viridisLite)
library(ggpubr)
library(dittoSeq)
library(utils)
library(SingleCellExperiment)
library(gridExtra)
library(Seurat)
library(cellhashR)
library(ggrepel)
library(scales)
library(demultiplexingbioc23)
```


# Benchmarking cell hashing based methods

Testing algorithms on real data is a key step to evaluating true performance.
Real data includes nuances and factors which may not have been foreseen when developing the algorithm.
The cost of scRNASeq means that few demultiplexing benchmarking datasets are available and often results from other demultiplexing algorithms are used as ground truth.
As demultiplexing performance is closely tied to data quality and hetergenity, the few benchmarking datasets may not give a representative account of algorithm performance across the range of data quality observed in practice.

To overcome this, with the added benefit of hindsight, we simulate data to examine algorithm performance under specific conditions associated with poor hashing quality.

## Simulating Data

````{r warning=FALSE}

mat<-logimat(ngroups=2,nsinglet=c(200,200),ndoub=50,nneg=50)
library(ComplexHeatmap)
mat[1:2,1:10]
Heatmap(mat)

counts<-draw_counts(size_sig=10,size_bg = 10,mu_sig = rep(50,2),mu_bg = rep(5,2),mat=mat,seed=1)
counts[1:2,c(1,201,401,451)]
Heatmap(log(counts+1))
df<-as.data.frame(t(counts))

myplots<-plot_hashtag(df)
do.call(grid.arrange,myplots)
````

## High quality hashing

````{r fig.width=9,fig.height=3}
set.seed(1)
mu1<-600

mat<-logimat(ngroups=4,nsinglet=c(2000,2000,2000,2000),ndoub=1000,nneg=200)
counts<-draw_counts(size_sig=10,size_bg = 3.5,mu_sig = c(mu1*1,mu1*1,mu1*1,mu1*1),mu_bg = rep(10,4),mat=mat,seed=1)
hto<-SingleCellExperiment(list(counts=counts))

seurat<-CreateSeuratObject(counts=counts,assay="HTO")
seurat<-NormalizeData(seurat,normalization.method ="CLR")
seurat$hashtag<-as.factor(colnames(counts))
seurat<-HTODemux(seurat)
#acc<-(sum(seurat$hashtag==seurat$hash.ID))/length(seurat$hashtag)
seurat$hashtag<-factor(seurat$hashtag,levels=c("Hashtag1","Hashtag2","Hashtag3","Hashtag4","Negative","Doublet"))
groups<-levels(seurat$hashtag)
precision<-c()
recall<-c()
for (i in seq_along(groups)) {

k<-groups[i]
cond<-seurat$hashtag==k
recall[i]<-sum(seurat$hashtag[cond]==seurat$hash.ID[cond])/length(seurat$hashtag[cond])
cond<-seurat$hash.ID==k
precision[i]<-sum(seurat$hashtag[cond]==seurat$hash.ID[cond])/length(seurat$hashtag[cond])

}
groups<-factor(groups,levels=c("Hashtag1","Hashtag2","Hashtag3","Hashtag4","Negative","Doublet"))
df<-data.frame(precision,recall,groups)

#split=seurat$hashtag=="Hashtag4" & seurat$hash.ID!= "Hashtag4"
myplots<-plot_hashtag(as.data.frame(t(seurat[["HTO"]]@data)),split=NULL)
#myplots[[5]]<-xy
do.call(grid.arrange,c(myplots,nrow=1))

````
````{r}

xy<-ggplot(df, aes(x=recall, y=precision,color=groups)) + 
  geom_point(show.legend = FALSE,size=3) + 
  scale_color_manual(values=dittoColors()[1:6]) + 
  geom_text_repel(aes(label=groups), size=3,show.legend = FALSE,force=10) +
  scale_x_continuous(labels = label_number(accuracy = 0.01)) +
  scale_y_continuous(labels = label_number(accuracy = 0.01)) + 
  xlim(0,1) +
  ylim(0,1) + ggtitle("HTODemux")
xy

````
## Poor quality hashing

````{r fig.width=9,fig.height=3,warning=FALSE}
set.seed(1)
mu1<-400
mu2<-20
mat<-logimat(ngroups=4,nsinglet=c(2000,1000,3200,150),ndoub=1000,nneg=200)
counts<-draw_counts(size_sig=2.5,size_bg = c(2.5,2.5,2,2.5),mu_sig = c(mu1*1.2,mu1*0.9,mu1*0.7,mu1*0.9),mu_bg = c(mu2,mu2,mu2,mu2),mat=mat,seed=1)
hto<-SingleCellExperiment(list(counts=counts))

seurat<-CreateSeuratObject(counts=counts,assay="HTO")
seurat<-NormalizeData(seurat,normalization.method ="CLR")
seurat$hashtag<-as.factor(colnames(counts))
seurat<-HTODemux(seurat)
#acc<-(sum(seurat$hashtag==seurat$hash.ID))/length(seurat$hashtag)
seurat$hashtag<-factor(seurat$hashtag,levels=c("Hashtag1","Hashtag2","Hashtag3","Hashtag4","Negative","Doublet"))
groups<-levels(seurat$hashtag)
precision<-c()
recall<-c()
for (i in seq_along(groups)) {

k<-groups[i]
cond<-seurat$hashtag==k
recall[i]<-sum(seurat$hashtag[cond]==seurat$hash.ID[cond])/length(seurat$hashtag[cond])
cond<-seurat$hash.ID==k
precision[i]<-sum(seurat$hashtag[cond]==seurat$hash.ID[cond])/length(seurat$hashtag[cond])

}
groups<-factor(groups,levels=c("Hashtag1","Hashtag2","Hashtag3","Hashtag4","Negative","Doublet"))
df<-data.frame(precision,recall,groups)

#split=seurat$hashtag=="Hashtag4" & seurat$hash.ID!= "Hashtag4"
myplots<-plot_hashtag(as.data.frame(t(seurat[["HTO"]]@data)),split=NULL,xlim=c(0,4),ylim=c(0,2000))
#myplots[[5]]<-xy
do.call(grid.arrange,c(myplots,nrow=1))

````

````{r}

xy<-ggplot(df, aes(x=recall, y=precision,color=groups)) + 
  geom_point(show.legend = FALSE,size=3) + 
  scale_color_manual(values=dittoColors()[1:6]) + 
  geom_text_repel(aes(label=groups), size=3,show.legend = FALSE,force=10) +
  scale_x_continuous(labels = label_number(accuracy = 0.01)) +
  scale_y_continuous(labels = label_number(accuracy = 0.01)) + 
  xlim(0,NA) +
  ylim(0,NA) + ggtitle("HTODemux")
xy

````

````{r}

df<-data.frame(true=seurat$hashtag,pred=seurat$hash.ID)
df$true<-factor(df$true,levels=c("Hashtag1","Hashtag2","Hashtag3","Hashtag4","Negative","Doublet"))
df$pred<-factor(df$pred,levels=c("Hashtag1","Hashtag2","Hashtag3","Hashtag4","Negative","Doublet"))

count<-ggplot(df, aes(fill=true,x=pred))+geom_bar() + scale_fill_manual(values=dittoColors())
prop<-ggplot(df, aes(fill=true,x=pred))+geom_bar(position="fill") + scale_fill_manual(values=dittoColors()) +ylab("proportion")
count+prop
````

# Session Info

```{r}
sessionInfo()
```

# References
